Graph Topological Aspects of Granger Causal Network Learning\\ \large Supplementary Material

### 2019-07-24 paper_theory/supplementary_material.tex
Some renaming and prepping for submission.

\documentclass{statsoc}
\usepackage{fullpage}
\usepackage{ccfonts,eulervm}
\usepackage{tabularx}
\usepackage{xr}
\usepackage{chngcntr}
\usepackage{apptools}
\AtAppendix{\counterwithin{lemma}{section}}
\AtAppendix{\counterwithin{definition}{section}}
\AtAppendix{\counterwithin{equation}{section}}
\externaldocument{paper_theory}

@@

  where $(a)$ applies the orthogonality principle, $(b)$ expands with
  Equation \eqref{eqn:ancestor_expansion} with
  $H_i(z)x_i(t) = \linE{x_i(t)}{H_{t - 1}^i}$, and $(c)$ follows by
  performing cancellations of $v_k(t) \perp v_\ell(t)$ since by the
  contrapositive of Proposition \ref{prop:sc_graph_common_anc} we
  cannot have $i \in \anc{j}$ or $j \in \anc{i}$.

@@

  \begin{align*}
    &\sum_{k \in \anc{i} \cap \anc{j}}\inner{F(z)\A_{jk}(z)v_k(t)}{(1 - H_i(z))\A_{ik}(z)v_k(t)}\\
    = &\sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))F(z)\A_{jk}(z)v_k(t)}{v_k(t)}\\
  \end{align*}

@@

  Since each $v_k(t)$ is uncorrelated through time,
  $\inner{T_{ij}(z)v_k(t)}{v_k(t)} = \sigma_k^2T_{ij}(0)$, and
  therefore we have $j \pwgc i$ if $T_{ij}(z)$ is \textit{not} causal
  and $i \pwgc j$ if $T_{ij}(z)$ it \textit{not} anti-causal.
  Moreover, we have $i \npwgc j$ \textit{and} $j \pwgc i$ if
  $T_{ij}(z)$ is a constant.  Therefore, under Assumption
  \ref{ass:T_causality} $j \pwgc i \implies i \pwgc j$.

@@

\begin{lemma}[$W$ Represents Ancestor Relations]
  \label{lem:W_subset_E}
  In Algorithm \ref{alg:pwgr} we have
  $(i, j) \in W$ if and only if $i \in \anc{j}$.  In particular,
  $W \subseteq \gcge$.
\end{lemma}
\begin{proof}
  Let $j \in [n]$ and suppose that $i \in \anc{j}$.  Then $i \pwgc j$
  by Proposition \ref{prop:pwgc_anc}.  Proposition
  \ref{prop:sc_graph_common_anc} ensures that $(i, j)$ are not
  confounded and Corollary \ref{cor:parent_corollary} that
  $j \not\in \anc{i}$ so $j \npwgc i$ and thence by Proposition
  \ref{prop:ancestor_properties} $(i, j) \in W$.

  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
  Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
  are not confounded and so by Proposition \ref{prop:ancestor_properties}
  we must have $i \in \anc{j}$.
\end{proof}

@@

\begin{lemma}[Depth Characterization of $P_k$]
  \label{lem:depth_lemma}
  $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
\end{lemma}
\begin{proof}
  We proceed by induction, noting that $P_0$ is non-empty since $\gcg$
  is acyclic and therefore $\gcg$ contains nodes without parents.  The
  base case $i \in P_0 \iff d(i) = 0$ is by definition, and
  $j \in S_0 \iff d(j) \ge 0$ is trivial since $S_0 = [n]$.  So
  suppose that the lemma is true up to $k - 1$.

@@

  ($d(i) = k \implies i \in P_k$): Suppose $i \in [n]$ is such that
  $d(i) = k$.  Then $i \in S_{k - 1}$ by the hypothesis, but also
  $i \not\in P_{k - 1}$ so then
  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$ and thus $d(i) \ge k$.
  Now, recalling the definition of $P_k$

@@

  if $s \in S_k$ is such that $(s, i) \in W$ then $s \pwgc i$ and
  $i \npwgc s$ so that by persistence and Proposition
  \ref{prop:persistence_converse} there cannot be a confounder of
  $(s, i)$ (otherwise $i \pwgc s$) so then by Proposition
  \ref{prop:ancestor_properties} we have $s \in \anc{i}$.  We have
  shown that $s \in S_k \implies d(s) \ge k$ and so we must have
  $d(i) > k$, a contradiction, thence $s \not\in \anc{i}$,
  $s \npwgc i$, $(s, i) \not\in W$ and $i \in P_k$.

@@

\begin{lemma}[Inner Loop]
  \label{lem:inner_loop_lemma}
  Fix an integer $k \ge 1$ and suppose that $(i, j) \in E_{k - 1}$ if
  and only if $(i, j) \in \gcge$ and $d(j) \le k - 1$.  Then, we have
  $(i, j) \in D_{kr}$ if and only if $(i, j) \in \gcge $, $d(j) = k$,
  and $d(i) = k - r$.
\end{lemma}
\begin{proof}
  We prove by induction on $r$, keeping in mind the results of Lemmas
  \ref{lem:W_subset_E} and \ref{lem:depth_lemma}.  For the base case,
  let $r = 1$ and suppose that $(i, j) \in \gcge$ with $d(j) = k$ and
  $d(i) = k - 1$.  Then, $(i, j) \in W$ and by our assumptions on
  $E_{k - 1}$ there is no $\gcgpath{i}{j}$ path in $E_{k - 1}$
  and therefore $(i, j) \in D_{k1}$.  Conversely, suppose that
  $(i, j) \in D_{k1}$.  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
  $(i, j) \in W \implies i \in \anc{j}$ implies that
  $i \in \pa{j}$ and $(i, j) \in \gcge$.

@@

  Conversely, suppose that $(i, j) \in \gcge$ and $d(j) \le k$.  If
  $d(j) < k$ then by induction $(i, j) \in E_{k - 1} \subseteq E_k$ so
  suppose further than $d(j) = k$.  Since $i \in \pa{j}$ we must have
  $d(i) < k$ (else $d(j) > k$) and again by Lemma
  \ref{lem:inner_loop_lemma} $(i, j) \in \bigcup_{r = 1}^k D_{kr}$
  which implies that $(i, j) \in E_k$.  The result follows.
\end{proof}

\begin{lemma}[Finite Termination]
  Algorithm \ref{alg:pwgr} terminates and returns $E_{k^\star - 1} = \gcge$
  for some $k^\star \le n$.
\end{lemma}
\begin{proof}
  If $n = 1$, the algorithm is clearly correct, returning on the first
  iteration with $E_1 = \emptyset$.  When $n > 1$ Lemma
  \ref{lem:outer_loop_lemma} ensures that $E_k$ coincides with
  $\{(i, j) \in \gcge\ |\ d(j) \le k\}$ and since $d(j) \le n - 1$ for
  any $j \in [n]$ there is some $k^\star \le n$ such that
  $E_{k^\star - 1} = \gcge$.  We must have $S_{k^\star} = \emptyset$
  since $j \in S_{k^\star} \iff d(j) \ge k^\star$ (if $d(j) > k - 1$ then
  $E_{k^\star - 1} \ne \gcge$) and therefore the algorithm terminates.
\end{proof}

@@

\begin{align}
  H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
  H_1:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
\end{align}

@@

\subsection{Edge Probabilities and Error Rate Controls}
\label{sec:error_rate_control}
Denote $F_{ij}$ the Granger-causality statistic of equation
\ref{eqn:gc_statistics} with model orders chosen by the methods of
Section \ref{sec:model_order_selection}.  We assume that this
statistic is asymptotically $\chi^2(p_{ij})$ distributed (the
disturbances are Gaussian), and denote by $G$ the cumulative
distribution function thereof.  We will define the matrix

@@

\clearpage
\bibliography{\string~/Documents/academics/global_academics/global_bib}

\end{document}

>>>>

\documentclass{statsoc}
\usepackage{fullpage}
\usepackage{ccfonts,eulervm}
\usepackage{tabularx}
\usepackage{xr}
\usepackage{chngcntr}
\usepackage{apptools}
\AtAppendix{\counterwithin{lemma}{section}}
\AtAppendix{\counterwithin{definition}{section}}
\AtAppendix{\counterwithin{equation}{section}}
\externaldocument{graph_topology_granger_causality}

@@

  where $(a)$ applies the orthogonality principle, $(b)$ expands with
  Equation \eqref{eqn:ancestor_expansion} with
  $H_i(z)x_i(t) = \linE{x_i(t)}{H_{t - 1}^i}$, and $(c)$ follows by
  performing cancellations of $v_k(t) \perp v_\ell(t)$ and noting that
  by the contrapositive of Proposition \ref{prop:sc_graph_common_anc}
  we cannot have $i \in \anc{j}$ or $j \in \anc{i}$.

@@

  \begin{align*}
    &\sum_{k \in \anc{i} \cap \anc{j}}\inner{F(z)\A_{jk}(z)v_k(t)}{(1 - H_i(z))\A_{ik}(z)v_k(t)}\\
    = &\sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))F(z)\A_{jk}(z)v_k(t)}{v_k(t)}.\\
  \end{align*}

@@

  Since each $v_k(t)$ is uncorrelated through time,
  $\inner{T_{ij}(z)v_k(t)}{v_k(t)} = \sigma_k^2T_{ij}(0)$, and
  therefore we have $j \pwgc i$ if $T_{ij}(z)$ is \textit{not} causal
  and $i \pwgc j$ if $T_{ij}(z)$ it \textit{not} anti-causal.
  Moreover, we have $i \npwgc j$ \textit{and} $j \pwgc i$ if
  $T_{ij}(z)$ is a constant.  Therefore, under Assumption
  \ref{ass:T_causality} $j \pwgc i \iff i \pwgc j$.

@@

\begin{lemma}[$W$ Represents Ancestor Relations]
  \label{lem:W_subset_E}
  In Algorithm \ref{alg:pwgr} we have
  $(i, j) \in W$ if and only if $i \in \anc{j}$.  In particular,
  $W \subseteq \gcge$.
\end{lemma}
\begin{proof}
  Let $j \in [n]$ and suppose that $i \in \anc{j}$.  Then $i \pwgc j$
  by Proposition \ref{prop:pwgc_anc}.  Proposition
  \ref{prop:sc_graph_common_anc} ensures that $(i, j)$ are not
  confounded and Corollary \ref{cor:parent_corollary} that
  $j \not\in \anc{i}$ so $j \npwgc i$ by Proposition and therefore
  \ref{prop:ancestor_properties} $(i, j) \in W$.

  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$,
  Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
  are not confounded and so by Proposition \ref{prop:ancestor_properties}
  we must have $i \in \anc{j}$.
\end{proof}

@@

\begin{lemma}[Depth Characterization of $P_k$ and $S_k$]
  \label{lem:depth_lemma}
  $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
\end{lemma}
\begin{proof}
  We proceed by induction, noting that $P_0$ is non-empty since $\gcg$
  is acyclic and therefore $\gcg$ contains nodes without parents.  The
  base case $i \in P_0 \iff d(i) = 0$ is by definition, and
  $j \in S_0 \iff d(j) \ge 0$ is trivial since $S_0 = [n]$.  So
  suppose that the lemma is true up to $k - 1$.

@@

  ($d(i) = k \implies i \in P_k$): Suppose $i \in [n]$ is such that
  $d(i) = k$.  Then $i \in S_{k - 1}$ by the hypothesis, but also
  $i \not\in P_{k - 1}$ so then
  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$.  Now, recalling the
  definition of $P_k$

@@

  if $s \in S_k$ is such that $(s, i) \in W$ then $s \pwgc i$ and
  $i \npwgc s$ so that by Proposition \ref{prop:persistence_converse}
  there cannot be a confounder of $(s, i)$ (otherwise $i \pwgc s$) so
  then by Proposition \ref{prop:ancestor_properties} we have
  $s \in \anc{i}$.  We have shown that $s \in S_k \implies d(s) \ge k$
  and so we must have $d(i) > k$, a contradiction, therefore there is
  no such $s \in S_k$ so $i \in P_k$.

@@

\begin{lemma}[Inner Loop]
  \label{lem:inner_loop_lemma}
  Fix an integer $k \ge 1$ and suppose that $(i, j) \in E_{k - 1}$ if
  and only if $(i, j) \in \gcge$ and $d(j) \le k - 1$.  Then, we have
  $(i, j) \in D_{kr}$ if and only if $(i, j) \in \gcge $, $d(j) = k$,
  and $d(i) = k - r$.
\end{lemma}
\begin{proof}
  We prove by induction on $r$, keeping in mind the results of Lemmas
  \ref{lem:W_subset_E} and \ref{lem:depth_lemma}.  For the base case,
  let $r = 1$ and suppose that $(i, j) \in \gcge$ with $d(j) = k$ and
  $d(i) = k - 1$.  Then, by Corollary \ref{cor:gc_implies_pwgc}
  $(i, j) \in W$ and by our assumptions on $E_{k - 1}$ there is no
  $\gcgpath{i}{j}$ path in $E_{k - 1}$ and therefore
  $(i, j) \in D_{k1}$.  Conversely, suppose that $(i, j) \in D_{k1}$.
  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
  $(i, j) \in W \implies i \in \anc{j}$ implies that $i \in \pa{j}$
  and $(i, j) \in \gcge$.

@@

  Conversely, suppose that $(i, j) \in \gcge$ and $d(j) \le k$.  If
  $d(j) < k$ then by induction $(i, j) \in E_{k - 1} \subseteq E_k$ so
  suppose further than $d(j) = k$.  Since $i \in \pa{j}$ we must have
  $d(i) < k$ (else $d(j) > k$) and again by Lemma
  \ref{lem:inner_loop_lemma} $(i, j) \in \bigcup_{r = 1}^k D_{kr}$
  which implies that $(i, j) \in E_k$.
\end{proof}

\begin{lemma}[Finite Termination]
  Algorithm \ref{alg:pwgr} terminates and returns the set
  $E_{k^\star - 1} = \gcge$ for some $k^\star \le n$.
\end{lemma}
\begin{proof}
  If $n = 1$, the algorithm is clearly correct, returning on the first
  iteration with $E_1 = \emptyset$.  When $n > 1$ Lemma
  \ref{lem:outer_loop_lemma} ensures that $E_k$ coincides with
  $\{(i, j) \in \gcge\ |\ d(j) \le k\}$ and since $d(j) \le n - 1$ for
  any $j \in [n]$ there is some $k^\star \le n$ such that
  $E_{k^\star - 1} = \gcge$.  We must have $S_{k^\star} = \emptyset$
  since $j \in S_{k^\star} \iff d(j) \ge k^\star$ (if $d(j) > k - 1$ then
  $E_{k^\star - 1} \ne \gcge$) and therefore the algorithm terminates.
\end{proof}

@@

\begin{align}
  H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
  H_1:&\ \widehat{x}_{i|j}^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
\end{align}

@@

\subsection{Edge Probabilities and Error Rate Controls}
\label{sec:error_rate_control}
Denote $F_{ij}$ the Granger-causality statistic of equation
\ref{eqn:gc_statistics} with model orders chosen by the methods of
Section \ref{sec:model_order_selection}.  We assume that this
statistic is asymptotically $\chi^2(p_{ij})$ distributed (e.g. the
disturbances are Gaussian), and denote by $G$ the cumulative
distribution function thereof.  We will define the matrix

@@

\clearpage
% \bibliography{\string~/Documents/academics/global_academics/global_bib}
\bibliography{paper_bib}
\end{document}
